---
title: "Species lists with abundance data and biodiversity metrics"
output: html_notebook
---


##Introduction - a simple example

I'm going to develop examples first shown in our Wednesday class on February 13th. 

I was creating example "plant communities" using simply a typed letter to indicate an individual plants. So "A A A A A A" represents six individuals of species A. Then "B B B B B B" represents six individuals of a new species, B. 

We can do this by creating a character variable, with multiple entries. 

```{r}
one=c("A","A","A","A","A","A","B","B","B","B","B","B","C","C","C","C","C","C","D","E","F","G","H")
one
length(one)
```

I asked for the variable length, which is 23. That tells us that there are 23 individuals in the sample. 


```{r}
#Now we want to figure out something about abundances within species. Let's figure out, simply by counting,  how many individuals there are of each species. This is a measure of the abundance of each species. For this, we use the function "count" which is in a package called PLYR, which we must first load. Once that's done,  we create a new variable which has the count of each species
#library(plyr)
one
one.ab=count(one)
one.ab
class(one.ab)
length(one.ab)
#We might also want to get a list of species, and the length of that list (which is species richness). 
one.u=unique(one)
one.u
length(one.u)
attributes(one.ab)
```

```{R}
#But before we do the combining, we must add a variable to allow keeping track of which sample is which
one.ab=data.frame("one", one.ab)
one.ab
#This added a first column with every row populated by the string variable "one"
#Let's also re-name columns. Later, our analyses of species richenss will need specifically named variables. 
names(one.ab)<-c("ListID", "SpeciesID", "Freq")
one.ab
```

Now let's do that for another sample:

```{r}
three=c("A","A","A","A","A","A","B","B","B","B","B","B","B","C","C","C","C","C","C","C","D","E","F","G","Q")
three.ab=count(three)
three.ab
#We might also want to get a list of species, and the length of that list (which is species richness). 
three.u=unique(three)
three.u
length(three.u)
```


```{R}
#We can combine two files and compare them. 
#Once again, let's add a variable to allow keeping track of which sample is which
three.ab=data.frame("three", three.ab)
three.ab
names(three.ab)<-c("ListID", "SpeciesID", "Freq")
#And then combine the two variables into one
ab=rbind(one.ab, three.ab)
ab
dim(ab)
#A powerful function in LABDSV converts this list of sample IDs, species and site-specific abundances into a matrix in which species names are column headings, and row names are samples. To do this, it creates a master list of all species in all samples. A matrix like this is useful for quantifying multiple aspects of biodiversity -- alpha and beta and gamma. 
#So be sure to load the LABDSV package
library(labdsv)
abundance<-matrify(ab)
abundance
#If we look at the dimensions of this matrix, what does it mean? What is 2? What is 9?
dim(abundance)
```


## Matrix Power! Get a diversity index

Shannon-Weaver Diversity Index (H), possibly familiar from ecology or conservation biology classes, examines each species for the proportion of individuals that it contributes to the total individuals of all species in the sample (Pi). 

If S is the total number of species in the transect (i.e., the species richness), 
Diversity,   H  =  - SUM OVER species 1 to species i(Pi)*(ln(Pi))
							(NOTE: "ln" is the natural logarithm and i refers to the ith species)


As an example, let's return to the data above, where N total = 23 individuals:

```{r}
one
length(one)
```


where N total = 23 individuals

Let Ni be the number of individuals of any given species, and Pi = Ni/Ntotal:

Species 	Ni	  Pi              ln(Pi)   Pi * ln(Pi)
A 	      6	    6/23 = 0.26     -1.34   	-0.35
B	        6	    6/23 = 0.26   	-1.34   	-0.35
C	        6	    6/23 = 0.26   	-1.34   	-0.35
D         1     1/23 = 0.045	  -3.13     -0.136
E         1     1/23 = 0.045	  -3.13     -0.136
F         1     1/23 = 0.045	  -3.13     -0.136
G         1     1/23 = 0.045	  -3.13     -0.136
H         1     1/23 = 0.045	  -3.13     -0.136
Q         0     0/23 = 0          0        0
				H = -(-1.73)
				
You can try doing a similar calculation for Sample "three" and then compare to the values obtained from R using the VEGAN package

```{r}
#library(vegan)
abundance
shannon<-diversity(abundance, index = "shannon")
shannon
```
				

Another measure of diversity, Simpson's index. There are actually a lot of different versions of Simpsons index, all roughly based on the likelihood that any two randomly encountered individuals are the same species. 

In the package vegan, it is calculated as

D = 1 - Sum(p^2)

where p = n/N

where each n refers to the number of organisms of that particular species
and N refers to the number of organisms of all species

Species	ni	p=n/N           p^2  
A	      6	  6/23=0.26       0.068
B	      6	  6/23=0.26       0.068
C	      6	  6/23=0.26       0.068
D	      1	  1/23=0.043      0.00189
E	      1	  1/23=0.043      0.00189
F	      1	  1/23=0.043      0.00189
G	      1	  1/23=0.043      0.00189
H	      1	  1/23=0.043      0.00189
Q       0	  0

                SUM         0.214

D = 1 - Sum(p^2) = 0.786

```{r}
abundance
simpson<-diversity(abundance, index = "simpson")
simpson
```




```{r}
#We need to add more samples if we want to move toward making multiple pairwise comparisons
two=c("A","A","A","A","A","A","B","B","B","B","B","B","B","B","C","C","C","C","C","C","D","D","D","D","D","D","D","E","F","E","E","E","E","E","F","F","F","F","F","F","F","F","F","G","G","G","G","G","G","H","H","H","H","H","H","H","H")
five=c("A","A","A","A","A","A","B","B","B","B","B","B","B","D","D","D","D","D","D","D","J","K","L","M","N")
#Here are two additional samples named "two" and "five"
#By adding to this chunk of code (and running it), see if you can add to the output above to obtain similar data
two.ab=count(two)
two.ab
#We might also want to get a list of species, and the length of that list (which is species richness). 
two.u=unique(two)
two.u
length(two.u)
five.ab=count(five)
five.ab
#We might also want to get a list of species, and the length of that list (which is species richness). 
five.u=unique(five)
five.u
length(five.u)
```


```{R}
#Let's add more data so we can continue to delve into what can be done
two.ab
#Repeating what we did earlier, to process these lists so that we can incorporate them into an abundance matrix
two.ab=data.frame("two", two.ab)
two.ab
five.ab=data.frame("five", five.ab)
five.ab
names(two.ab)<-c("ListID", "SpeciesID", "Freq")
names(five.ab)<-c("ListID", "SpeciesID", "Freq")
one.ab
three.ab
two.ab
five.ab
ab=rbind(one.ab, three.ab, two.ab, five.ab)
ab
dim(ab)
#ab<-ab[,2:4]
abundance<-matrify(ab)
abundance
```



As mentioned before, this type of abundance matrix combined with PACKAGE VEGAN is powerful.

What can we do with this bigger matrix? For starters, we can use an index to measure the similarity (or dissimilarity) of any two pairs of lists in the matrix

There are many different ways to measure community similarity or dissimilarity. 

An enduringly useful pairwise index that uses abundance data is called the Bray-Curtis coefficient, which is a ratio. 

First, to get the top/numerator of the ratio: for all species across all samples, compute the difference in abundance between a pair of samples (as an absolute value). Then, all of those absolute differences are summed.

Next, to get the bottom/denominator of the ratio: sum the total individuals in the pair of samples

For our simple examples of samples "one" and "five": 

            "one"           "five"      |abs diff|
A 	           6	            6               0
B	             6	            7	              1 
C              6              0               6
D	             1              7	              6
E		           1              0               1
F		           1              0               1 
G		           1              0               1 
H		           1              0               1 
Q		           0              0               0 
J		           0              1               1
K		           0              1               1
L		           0              1               1
M		           0              1               1
N		           0              1               1

sum           23             25              22


Bray-Curtis = 22/(23+25) = 0.458


A simple pairwise index that uses only presence-absence data is called Jaccard's Coefficient of similarity, J. 

J = a / (n - d)

where a is the number of "positive matches" (+ +, species found in two lists), d is the number of negative matches (- - , species found in neither list) and n is the total number of possible matches (based on the total number of taxa found across all lists). 

For our simple examples of samples "one" and "five": 

            "one"           "five" 
A 	           x	            x
B	             x	            x	 
C              x
D	             x              x	
E		           x
F	             x  	
G	             x	
H	             x
Q	           	
J		                          x
K                             x
L                             x
M                             x    
N                             x


in which a = 3, d = 1, n = 14

so J  =  3 / (14 - 1) = 0.769 

This index will vary between zero and one. 

There are many other ways to measure pair-wise similarities. 

If there are more than two lists, it's possible to put together a matrix of pair-wise values for the index, which is what the code below does. 

```{r}
#the function vegdist is used on the matrix, the metric specified. The assumption is that the data are abundance data, even though some indices require binary data. So the unstated and default argument that is always used in vegdist is that  binary=FALSE. This can and should be changed if computing a presence-absence metric like Jaccard.
bray = vegdist(abundance, "bray") 
bray
jaccard = vegdist(abundance, "jaccard", binary=T)
jaccard
```



```{r}
#Now that there are four samples, it makes sense to start seeing if this is enough samples to gauge diversity across these samples. Vegan can compute the cumulative number of species across all four, and then plot them. It will take the plots in random order. Other options for this method argument include "collector" which is the order in which they are entered into the matrix. Another is "rarefaction" which will adjust to take into account that each sample may have a different number of individuals.
accurve<-specaccum(abundance, method="random", permutations=100)
plot(accurve$sites, accurve$richness,
     xlab="Number of Sites",
     ylab="Species Richness",
     main="Now with four samples, random order")
accurve2<-specaccum(abundance, method="collector", permutations=100)
plot(accurve2$sites, accurve$richness,
     xlab="Number of Sites",
     ylab="Species Richness",
     main="As encountered, with four samples")
```


## Multi-variate statistics for abundance data

#### NMDS Ordination (non-metric multi-dimensional scale)

This is some advanced stuff in terms of statistics, but what's powerful and fairly intuitive are the plots that you can get using this NMDS analysis which uses the vegan function called metaMDS.

```{r}
head(abundance)
example_NMDS=metaMDS(abundance, k=3) # Our community-by-species matrix 
# K=2 is setting the number of reduced dimensions. Increase if high stress is problem. 
#"The stress, or the disagreement between 2-D configuration and predicted values from the regression"

#A good rule of thumb: stress > 0.05 provides an excellent representation in reduced dimensions, > 0.1 is great, >0.2 is good/ok, and stress > 0.3 provides a poor representation

plot(example_NMDS)
test.plot<-ordiplot(example_NMDS,type="n") #Ordination plot function especially for congested plots
points(test.plot, "species", pch=21, col="red", bg="yellow")

test.plot<-ordiplot(example_NMDS,type="n") #Ordination plot function especially for congested plots
points(test.plot, "sites", col="blue", cex=0.9)

test.plot<-ordiplot(example_NMDS,type="n") #Ordination plot function especially for congested plots
points(test.plot, "species", pch=21, col="red", bg="yellow")
text(test.plot, "sites", col="blue", cex=0.9)

test.plot<-ordiplot(example_NMDS,type="n") #Ordination plot function especially for congested plots
text(test.plot, "species", pch=21, col="red", bg="yellow")
points(test.plot, "sites", col="blue", cex=0.9)


```

If all of the samples were the same, then this analysis would wind up finding that all of them measure near zero (the mean) on the x-axis, labeled as NMDS1, and likewise on NMDS2. These are on an arbitrary scale, with zero being the overall average. So samples one and three are different from that average in one direction, two is close to that average, and five is larger than that average. 

On the y-axis, two deviates in a different direction from one, three and five. 

In other words, two or more samples with a similar species composition and species abundance will wind up clustering near each other. Samples that differ in species composition tend to be spread far away from each other. 

## Switching to real species data: County occurrence data from BIEN

```{r}
#library("bien")
#As you know, you can obtain species occurrence data from the BIEN database, or from another even more comprehensive data base called gbif (gbif.org). Occurrence data is a type of abundance data, and it can also be used to figure out presence-absence data. 
bronx <- BIEN_occurrence_county(country="United States", state="New York", county = "Bronx")
head(bronx)
tail(bronx)
manhattan <- BIEN_occurrence_county(country="United States", state="New York", county = "New York")
head(manhattan)
tail(manhattan)
attributes(manhattan)
westchester <- BIEN_occurrence_county(country="United States", state="New York", county = "Westchester")
head(westchester)
tail(westchester)
attributes(westchester)
```

##Now I'm going to work with county data

```{r}
#library(BIEN)
#We previously learned to "slim down" data sets to one or a few variables, which we can do here
bronx.occ=bronx[c(1)]
head(bronx.occ)
manhattan.occ=manhattan[c(1)]
head(manhattan.occ)
westchester.occ=westchester[c(1)]
head(westchester.occ)
#These are now two species occurrence lists, one for The Bronx, one for Manhattan
#To turn them into species lists with abundance data, you can recycle the code from above.
bronx.ab=count(bronx.occ)
head(bronx.ab)
manhattan.ab=count(manhattan.occ)
head(manhattan.ab)
westchester.ab=count(westchester.occ)
head(westchester.ab)
#Then re-name the columns as previously. Here just the Bronx
bronx.ab=data.frame("bronx", bronx.ab)
head(bronx.ab)
names(bronx.ab)<-c("ListID", "SpeciesID", "Freq")
head(bronx.ab)
manhattan.ab=data.frame("manhattan", manhattan.ab)
head(manhattan.ab)
names(manhattan.ab)<-c("ListID", "SpeciesID", "Freq")
head(manhattan.ab)
westchester.ab=data.frame("westchester", westchester.ab)
head(westchester.ab)
names(westchester.ab)<-c("ListID", "SpeciesID", "Freq")
head(westchester.ab)
threecounty.ab=rbind(bronx.ab, manhattan.ab, westchester.ab)
head(threecounty.ab)
tail(threecounty.ab)
dim(threecounty.ab)
#This powerful function matrify is going to convert this list of species with their site-specific abundances into a matrix with the species as the column headings, the rows as the samples. This gives a long sample list
urban.ab<-matrify(threecounty.ab)
urban.ab
#If we look at the dimensions of this matrix, what does it mean? What is 2? What is 9?
dim(urban.ab)
simpson<-diversity(urban.ab, index = "simpson")
shannon<-diversity(urban.ab, index = "shannon")
simpson
shannon
bray = vegdist(urban.ab, "bray")
bray
jaccard = vegdist(urban.ab, "jaccard", binary=T)
jaccard


```

```{r}
write.csv(threecounty.ab, file = "threecounty.csv")
```

What if you wanted to save the abundance matrix as well?


```{r}
write.csv(urban.ab, file = "urban.csv")
```


Now that we know how to write a file in CSV format, how would we read it?

```{r}
#How about if you have another file? 
MH=read.csv("WelikiaMH.csv")
head(MH)
tail(MH)
```

```{r}
#Since the genus name and species names were in separate columns, I'm going to paste them back together
speciesID=paste(MH$Scientific, MH$Name, sep=" ")
speciesID
```

```{r}
#Just as we did before, let's count how often each species is in the occurrence list.
MH_list=count(speciesID)
head(MH_list)
MH_list=data.frame(MH_list)
MH_list=data.frame("MH", MH_list)
head(MH_list)
```
```{r}
#Now I'm going to give them headings, just as we did earlier
names(MH_list)<-c("ListID", "SpeciesID", "Freq")
```

```{r}
#Now I'm going to repeat that for my second sample
LES=read.csv("Welikiales.csv")
head(LES)
tail(LES)
speciesID=paste(LES$Scientific, LES$Name, sep=" ")
speciesID
LES_list=count(speciesID)
head(LES_list)
LES_list=data.frame("LES", LES_list)
names(LES_list)<-c("ListID", "SpeciesID", "Freq")
```

```{r}
#And if I combine my two lists into one
Welikia2=rbind(LES_list, MH_list)
dim(Welikia2)
#And run the matrify function
wab=matrify(Welikia2)
wab
```

```{r}
#Then I am ready to do some computing of an index or of several indices:
simpson<-diversity(wab, index = "simpson")
shannon<-diversity(wab, index = "shannon")
simpson
shannon
jaccard = vegdist(wab, "jaccard")
jaccard
bray = vegdist(wab, "bray")
bray
```





## How to finish this portfolio

The last step in finishing this portfolio is to prepare two types of lists. 

a. lists of three counties
b. some other type of list - ideally get three lists of the same type that are interesting to you (look back to Portfolio 2.1 on Courseworks)

Process your county lists until they are transformed into a species abundance matrix. 

Then compute some diversity indices. 
Before finalizing and turning in our notebook, add  commentary.


## a.Three counties
```{r}
Kings <- BIEN_occurrence_county(country="United States", state="New York", county = "Kings")
head(brooklyn)
tail(brooklyn)
Manhattan <- BIEN_occurrence_county(country="United States", state="New York", county = "New York")
head(Manhattan)
tail(Manhattan)
attributes(Manhattan)
Queens <- BIEN_occurrence_county(country="United States", state="New York", county = "Queens")
head(Queens)
tail(Queens)
attributes(Queens)
```

```{r}
manhattan.occ=Manhattan[c(1)]
head(manhattan.occ)

#ISSUE
queens.occ=Queens[c(1)]
head(queens.occ)

kings.occ=Kings[c(1)]
head(kings.occ)
```

```{r}
manhattan.ab=count(manhattan.occ)
head(manhattan.ab)
queens.ab=count(queens.occ)
kings.ab=count(kings.occ)

```

```{r}

manhattan.ab=data.frame("Manhattan", manhattan.ab)
names(manhattan.ab)<-c("ListID", "SpeciesID", "Freq")
head(manhattan.ab)

queens.ab=data.frame("Queens", queens.ab)
names(queens.ab)<-c("ListID", "SpeciesID", "Freq")
head(queens.ab)

kings.ab=data.frame("Kings", kings.ab)
names(kings.ab)<-c("ListID", "SpeciesID", "Freq")
head(queens.ab)
```


#All three bound together
```{r}
tricounty.ab=rbind(manhattan.ab, queens.ab, kings.ab)
head(tricounty.ab)
tail(tricounty.ab)
dim(tricounty.ab)
threecounties.ab<-matrify(tricounty.ab)
head(threecounties.ab)

dim(threecounties.ab)
```


#diversity indexes of counties

#Shannon diversity index, bray, and jaccard/binary on
```{r}
brooklyn<-diversity(threecounties.ab, index = "shannon")
#brooklyn<-diversity(counties.ab, index = "Kings")
brooklyn

#PROBLEM? brooklyn = vegdist(threecounties.ab, "Kings")

brooklyn
bray = vegdist(threecounties.ab, "bray")
bray

jaccard = vegdist(threecounties.ab, "jaccard", binary=T)
jaccard
```

```{r}
threecounties.ab
shannon<-diversity(threecounties.ab, index = "shannon")
shannon
```


```{r}
threecounties.ab
simpson<-diversity(threecounties.ab, index = "simpson")
simpson
```

#I don't understand exactly what I'm looking at below. It doesn't seem to match the data?
```{r}
#cumulative number of species across all three

plotthreecounties<-specaccum(threecounties.ab, method="random", permutations=100)
plot(plotthreecounties$sites, plotthreecounties$richness,
     xlab="Number of Sites",
     ylab="Species Richness",
     main="Manhattan, Queens, and Kings Counties")
plotthreecounties2<-specaccum(threecounties.ab, method="collector", permutations=100)
plot(plotthreecounties2$sites, plotthreecounties$richness,
     xlab="Number of Sites",
     ylab="Species Richness",
     main="Manhattan, Queens and Kings, Counties: Rareification")
```

#Below is not really working out
```{r}
head(threecounties.ab)
threecountiestime=metaMDS(threecounties.ab, distance="bray", k=2) 

#This looks the same as the example code, so I must not have been setting things up correctly? Or not enough data? Stress seems to be at 0.0

plot(threecountiestime)
threecountiestime.plot<-ordiplot(threecountiestime,type="n") 
points(threecountiestime.plot, "species", pch=21, col="green", bg="light blue")
```

```{r}
threecountiestime.plot<-ordiplot(threecountiestime,type="n") #Ordination plot function especially for congested plots
points(threecountiestime.plot, "sites", col="red", cex=0.9)

threecountiestime.plot<-ordiplot(threecountiestime,type="n") #Ordination plot function especially for congested plots
points(test.plot, "species", pch=21, col="red", bg="yellow")
text(test.plot, "sites", col="blue", cex=0.9)

threecountiestime.plot<-ordiplot(threecountiestime,type="n") #Ordination plot function especially for congested plots
text(threecountiestime.plot, "species", pch=21, col="red", bg="yellow")
points(threecountiestime.plot, "sites", col="green", cex=0.9)
```



#How do you interpret the indices computed for the County data?

It looks like I may have misunderstood how to use these; it looks like I don't have enough data for the Counties in some instances. 


#The shannon index was 4.204 for Manhattan, 3.576 for Queens, and 3.422 for Kings (estimated to three decimals). So from what I've read, a high shannon index would mean high uncertainty, which would mean a higher diversity. These each have similar diversity, though Manhattan apparently has more diversity. BUT these are also hitting or exceeding the limits of the index, no? So I feel like I must have done something terribly wrong. Maybe I was not supposed to bind the datasets beforehand (threecounties.ab)?

No, these values don't exceed the limits of the index. You did it right! The Simpson index (below) is the one limited from 0 to 1. For both Shannon and Simpson indices, higher values = great diversity (i.e. species richness AND evenness).

#The simpson index approached 1; does this mean that each country is just about as diverse as the others? Yet, 1 should be infinite diversity, which doesn't really seem right for these areas. I'm unsure whether I should think of this as a ratio between each sample in a sort of circle/enclosed system that we are looking at with these ratios, but it's my first thought to do so. I also read that this is a measure of dominance, though, and that if it is high like this, it indicates low diversity, so maybe that's really how I should conceptualize this? I also read that this index is not linear, making the implications of .000'th decimals a bit more relevant. 


#The bray index is approaching 1 for all three counties. I think, as far as I know right now, that this means the abundance in each sample are very similar to one another, I think. If this is so, I now wish I had used counties that were not local at all to do a better comparison, and will play with this on my own (and also try to check whether I did something off to keep getting near 1:1 ratios)

No, with Bray-Curtis index, the 0 means less different, and 1 means more different. So if the values are close to 1 then your samples are not very similar. 

#I'm also unsure how to understand the way the jaccard index comes out, unless I did something wrong. It looks like it's saying that both Manhattan and Queens are .9912793 in some respect, Manhattan and Kings are both .9902913 in another respect, and Queens and Kings are both .9726027 in another respect; part of why I was thinking of the math being used as a closed system between three data sets. But, each of these are approaching the limit, which is 1, meaning the ratio is approaching 1. Does this then mean the ratio of present to absent species are about equal in each, and so each county has much overlap in species? Or does it mean the ratio indicates that each are just missing the same amount of species? I was trying to study up, but this is new territory for me completely (I’d never heard of any of this before), so it was hard to get it all in.

I agree. To be honest, after reading about them a bunch, Bray-Curtis and Jaccard are still confusing me, so we should talk to Dr. Callahan about this.

-------


## b.Three plant lists - 


```{r}
setwd("/Users/Natalie/Desktop/portfolio-2-eilatanz-MANUAL/")
planto=read.csv("04 Plant-o-matic Brooklyn.csv")
head(planto)
tail(planto)
names(planto)<-c("ListID", "SpeciesID", "Freq")
speciesID=paste(planto$Genus, planto$Species, sep=" ")
speciesID

```

```{r}
setwd("/Users/Natalie/Desktop/portfolio-2-eilatanz-MANUAL/")
queenspersist=read.csv("01 Science publication persisting plants in queens CLEANED.csv")

queenspersist
#stringsAsFactors = FALSE
head(queenspersist)
tail(queenspersist)
speciesID=(queenspersist$Species)
speciesID
```

```{r}
setwd("/Users/Natalie/Desktop/portfolio-2-eilatanz-MANUAL/")
chelsea=read.csv("02 Welikia Project 24 West 9th street Chelsea Manhattan 1609.csv")

#stringsAsFactors = FALSE removed from above, not needed
head(chelsea)
tail(chelsea)
speciesID=paste(chelsea$Genus, chelsea$Species, sep=" ")
speciesID

```

```{r}

chelsea_list=count(speciesID)
head(chelsea_list)
#df<-data.frame(words = unlist(words))
chelsea_list=data.frame(chelsea_list)
chelsea_list=data.frame("chelsea", chelsea_list)
names(chelsea_list)<-c("ListID", "SpeciesID", "Freq")

head(chelsea_list)
speciesID

queenspersist_list=count(speciesID)
head(queenspersist_list)
#df<-data.frame(words = unlist(words))
queenspersist_list=data.frame("queenspersist", queenspersist_list)
names(queenspersist_list)<-c("ListID", "SpeciesID", "Freq")

head(queenspersist_list)
speciesID

planto_list=count(speciesID)
head(planto_list)
planto_list=data.frame(planto_list)
planto_list=data.frame("planto", planto_list)
names(planto_list)<-c("ListID", "SpeciesID", "Freq")

head(planto_list)
speciesID
```

```{r}
#Run each of these separately for some reason...
chelsea_list
queenspersist_list
planto_list
#combine the lists
ThenNow=rbind(planto_list, chelsea_list)
dim(ThenNow)
#matrify function
plantochel=matrify(ThenNow)
plantochel
```
#how might I (and should I) give more rows with fewer columns/flip the axis on the above?

No, the whole point of the matrify function is to structure your data with a few rows and a lot of columns. Rows = sample areas, Columns = Unique species

#I tried to figure out why I could not bind all three using this method, and tried looking up how to add another column to the queens list (which is the problem apparently), but can't find out how. What am I doing wrong? Or am I using this for the wrong purpose? I did bind all three below in part b, though.

I think this wasn't working because previously on line 670 (I changed it) you had given the planto variable the wrong name, so your data.frame wasn't modified to have the correct column names. In order to bind three different data frames, they HAVE to have the same column names. So since your planto_list hadn't had the column names renamed correctly, it couldn't bind with the others.

```{r}
ThenNow2=rbind(planto_list, chelsea_list, queenspersist_list)
dim(ThenNow2)
#matrify function
plantochelquee=matrify(ThenNow2)
plantochelquee
```


#diversity indices


#Simpson index
```{r}
simpson<-diversity(plantochel, index = "simpson")
shannon<-diversity(plantochel, index = "shannon")
simpson
shannon
jaccard = vegdist(plantochel, "jaccard")
jaccard
bray = vegdist(plantochel, "bray")
bray
```


#lengths
```{r}
planto
length(planto)

chelsea
length(chelsea)

queenspersist
length(queenspersist)
```


#adding data, maybe I can do this better using below with species richness? I have a hard time understanding exactly what I'm calling from the server/databases with some of the commands, I think.
```{r}
planto.ab=count(planto)
planto.ab
 
planto.u=unique(planto)
planto.u
length(planto.u)


chelsea.ab=count(chelsea)
chelsea.ab

chelsea.u=unique(chelsea)
chelsea.u
length(chelsea.u)

queenspersist.ab=count(chelsea)
queenspersist.ab

queenspersist.u=unique(queenspersist)
queenspersist.u
length(queenspersist.u)

```
#adding more data

```{r}

#add third list 
planto.ab
planto.ab=data.frame("planto", planto.ab)
planto.ab

chelsea.ab=data.frame("chelsea", chelsea.ab)
chelsea.ab

queenspersist.ab=data.frame("queenspersist", queenspersist.ab)
queenspersist.ab

names(planto.ab)<-c("ListID", "SpeciesID", "Freq")
names(chelsea.ab)<-c("ListID", "SpeciesID", "Freq")
names(queenspersist.ab)<-c("ListID", "SpeciesID", "Freq")

#binding all three worked here, though:
abc=rbind(queenspersist.ab, planto.ab, chelsea.ab)
abc
dim(abc)
abc<-abc[,2:4]
lotto<-matrify(abc)
lotto
```


```{r}
 lottorandom<-specaccum(lotto, method="random", permutations=100)
plot(lottorandom$sites, lottorandom$richness,
     xlab="Number of Sites",
     ylab="Species Richness",
     main="Now with three samples, random order")
lottocollect<-specaccum(lotto, method="collector", permutations=100)
plot(lottocollect$sites, lottorandom$richness,
     xlab="Number of Sites",
     ylab="Species Richness",
     main="As encountered, with three samples")
```

```{r}
#library(vegan)
lotto
shannonlotto<-diversity(lotto, index = "shannon")
shannonlotto
```

```{r}
lotto
simpsonlotto<-diversity(lotto, index = "simpson")
simpsonlotto
```

```{r}
simpsonlotto<-diversity(lotto, index = "simpson")
shannonlotto<-diversity(lotto, index = "shannon")
simpsonlotto
shannonlotto
jaccardlotto = vegdist(lotto, "jaccard")
jaccardlotto
braylotto = vegdist(lotto, "bray")
braylotto
```


#I'm not clear on how to read the below output. Am I getting a bray index for each species across all three samples?
```{r}
braylotto = vegdist(lotto, "bray") 
braylotto
```

```{r}
jaccardlotto = vegdist(lotto, "jaccard", binary=T)
jaccardlotto
```

 
#How do you interpret indices computed from your second type of data? 

#So, in the lists of two bray and jaccard got 0, simpson and shannon each approached or got to the limit (.99 and 3.99); this occurred over and over. In the list of three, bray and jacquard each got 1, I think? As before, I think I must have some issue with my data or my code, and since this happened before, I’m guessing it’s my code? I’m not familiar enough with the concepts to feel like this is actually accurate, but it seems not to be to me so far. 

#The species richness plot I made was… not linear, but had quite the vector with the three datasets/samples. The resulting plot was also exactly the same as with random order. I think I must not be grasping what I’m calling down and why when I’m getting the data, or else something is wrong with my code. I need to spend more time with this I think, especially over break, if I'm to use it for a project. And I'd like to come to office hours for some help if I can! I feel a bit stunted, with the conceptual stuff and the coding being very new to me.

Yeah, there's a lot to go through here. I think you're making it more complicated then it needs to be, so let's find a time to meet to go through it from square one again. There are still some things I don't understand about a couple of the indices, so we can learn together! Sorry that I don't have more helpful comments at the moment. I just think working through this in person would be much more effective.